---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
data <- read.table(unz('activity.zip', 'activity.csv'),
                   header=TRUE, sep=",", stringsAsFactors=F,
                   colClasses=c('integer', 'Date', 'integer'))
```



## What is mean total number of steps taken per day?

```{r}
steps.per.day <- tapply(data$steps, data$date, sum, na.rm=T)
hist(steps.per.day, main='total number of steps taken each day', xlab='steps per day')
total.mean <- mean(steps.per.day, na.rm =T)
total.median <- median(steps.per.day, na.rm =T)
```

The mean total number of steps taken per day is `r sprintf("%8.2f", total.mean)`.

The mediean total number of steps taken per day is `r total.median`.



## What is the average daily activity pattern?

```{r}
steps.per.interval <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=T)
plot(steps.per.interval$x, type='l')
max.interval <- which.max(steps.per.interval$x)
```

The 5-minute interval with maximum number of steps on average across all the days in the dataset is on `r max.interval`-th 5-minute interval with `r round(steps.per.interval$x[max.interval], 2)` steps.



## Imputing missing values

```{r}
max.value.count <- sum(is.na(data$steps))
```

The total number of missing values in the dataset (i.e. the total number of rows with NAs) is `r max.value.count`.

Using mean for that 5-minute interval to fill all of the missing values.

```{r}
rounded.value.for.interval <- function(interval) {
  round(steps.per.interval$x[steps.per.interval$Group.1 == interval])
}

data.tidy <- data
data.tidy$steps[is.na(data$steps)] <- sapply(data$interval[is.na(data$steps)], rounded.value.for.interval)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
steps.per.day.tidy <- aggregate(c(data.tidy$steps) , by=list(data.tidy$date), FUN=sum)
names(steps.per.day.tidy) <- c('date', 'steps')
hist(steps.per.day.tidy$steps, main='total number of steps taken each day', xlab='steps per day')
tidy.mean <- mean(steps.per.day.tidy$steps)
tidy.median <- median(steps.per.day.tidy$steps)
```

The mean total number of steps taken per day is `r sprintf("%8.2f", tidy.mean)`.

The median total number of steps taken per day is `r sprintf("%8.0f", tidy.median)`.


```{r}
data.tidy$weekdays <- factor(c(
  ifelse(weekdays(data.tidy$date) == 'Sunday' | weekdays(data.tidy$date) == 'Saturday', 'weekend', 'weekday') ))

steps.per.weekdays <- aggregate(data.tidy$steps, by=list(data.tidy$interval, data.tidy$weekdays), FUN=mean)

names(steps.per.weekdays) <- c('interval','weekdays', 'steps')

library(lattice)
xyplot(steps ~ interval | weekdays, data=steps.per.weekdays, layout=c(1, 2), type='l')

```

